Transient Fault Tolerant Asynchronous Circuits

### 2019-12-06 draft_final.tex
before ch 3, 4 DD edits

    \title{Fault-tolerant Asynchronous Circuits}
    \author{Zoey Zhou}
    \dept{Electrical Engineering}
    \principaladviser{David Dill}
    \firstreader{Subhasish Mitra}
    \secondreader{Clark Barrett\\(Computer Science)}

@@

\prefacesection{Abstract}

\newpage

%% one can also have a prefacesection that is a Preface instead of
%% Acknowledgements.   The thematic purpose is the same (thanks).
    \prefacesection{Acknowledgements}
        I would like to thank Professor David Dill.

%% afterpreface produces a table of contents and any other tables
%% wanted. At the end pagenumbering changes from roman to arabic and
%% is restarted
    \afterpreface

@@

However, because the signals can arrive at the input at any time, the design is more complicated to ensure that the entire circuit functions correctly.  From a high level view the circuits are designed so that the signals progress irrespective of the delays.  This property is called Delay Insensitivity or Speed Independence depending on the underlying delay assumptions such as the maximum length of delays and on which components.  Even if this is designed correctly, when implemented there can still be hazards.  A hazard is when a certain sequence of delays of the circuit cause unwanted signal transitions to occur.  This may then propogate to other parts of the circuits and cause unwanted behavior.  When trying to eliminate hazards, the underlying delay assumptions are important as well as the implementation details such as gates and transistors. Given the same design under one implementation the circuit may have hazards while under another implemention the circuit will be hazard-free.  Some of the different delay assumptions that could be made are that the delays only occur at gates, or delays only occur on the wires or the delays occur on both gates and wires \cite{myers_book_2004}. Furthermore the bounds on the delay can vary as well. The most stringent case is if we assume that the delay can be arbitrarily long and takes on a value anywhere in the interval $(0,\infty)$ \cite{myers_book_2004}.  Circuits built under the unbounded gates and wires delay model are called Delay Insensitive circuits.  We will not use this delay model because the class of Delay Insensitive circuits is very restricted.  In particular one can only use buffers, inverters, and Muller C-elements to build Delay Insensitive circuits \cite{Martin_1990_DI} \cite{Martin1986_DI}.\\ %2nd reference not necessary

The asynchronous delay model that we consider is the unbounded gate delay model.  Circuits that can handle any delay under this model are called Speed Independent circuits. The unbounded gate delay model assumes that delays only appear on the gates and the delays can be arbitrarily long in the interval $(0,\infty)$.  The wire delays are negligible.  In addition we assume all forks, which is splitting a signal into multiple branches, do not have any time delays between each of its branches.  These assumptions are reasonable because the delays on the wires only change the operation of a circuit if the wire has forks.  (Otherwise you can lump the wire delay with the previous gate)  When a wire has forks, it is usually true in implementation that the delay between different branches of the fork is less than the delay for gates so that they arrive at the next gates at the same time from the perspective of the relevant gates.  This is also termed isochronic forks.  We focus on the unbounded gate delay model because the delay being unbounded is a pretty harsh requirement and would allow more circuits to work in a real life setting. \\ %how if we assume both gates and wires have delay cannot create many SI circuits, and also Muller is one of the broader classes that we can use

@@

The standard C-implementation is a gate-level implementation.  Each wire from the specification is implemented as a C-element with non-looping cascades of gates as the C-element's two inputs.  That is, each of the two inputs to the C-element is itself an output of a cascade of gates.  The C-element operates like a typical C-element where if the two inputs are `1', then the output is `1'; and when the two inputs are `0', the output is `0'; and when the two inputs have different values, the output value is held.  Additionally, Myers shows a method where the gate logic for the inputs to the C-element can always be re-implemented with two-input gates \cite{myers_book_2004}.  Since the C-element itself is also a two-input circuit element, then the entire circuit is composed of two-input circuit elements.  In such an implementation, each gate or C-element is assumed to have its own delay that is arbitrary in length.  Thus, this gate-level implementation allows for more generic design.
%explain sop?

\subsection{Quasi Delay Insensitive Circuits}
%QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits. Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays. In particular each branch of a fork will have different delays and can be treated as separate wires. One result about Delay Insensitive circuits is that it is a very restricted class. Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed. Note that in the case where all forks are isochronic, then this becomes the same as speed independent circuits. \\
The gates of a QDI circuit is formed from Production Rules (PRs). For the single output case, a gate takes a set of variables as inputs and outputs a variable. Again, a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$. Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards. Each guard is a boolean condition formed by some combination of the input variables. When a guard is evaluated to be $True$ then the transition is activated. When it is $False$ then the corresponding transition does not happen. Note that similar to the generalized C-elements, both guards cannot be active at the same time. In addition, it is assumed that each guard is in a sum of products of the input variables. The PRs of a circuit are evaluated all at the same time. At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$. This is due to the assumption of the arbitrary delay of a gate. Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allows any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

Another important property of QDI circuits is \textit{stability}. A guard is stable if it evaluates to False after the transition of the output variable is complete. This eliminates any hazards that may arise when a guard is True and before the output has fully transitioned, the guard then evaluates to False and the transition is terminated. This may cause a blip in the voltage, an incomplete transition, where the circuit may interpret it as either no transition or two consecutive transitions. One additional assumption is that the PRs cannot be \textit{self-invalidating}. That is, when a transition on an output occurs, it cannot immediately falsify the guard. With this assumption, it can be shown that the output of a gate cannot also be its own input variable. There are no self-loops.

\section{Fault Tolerant solutions}

%soft error protection <- error correcting codes for SRAM memory
Because faults cannot be eliminated in digital circuits, different methodologies have been used to make the circuits fault-tolerant. Much of the research done in this area are for synchronous circuits. There has also been some research for asynchronous circuits using some similar techniques. First, there is an overview of some common techniques for synchronous circuits, then there is a summary of the current techniques for asynchronous circuits. Some of the fault-tolerant techniques in asynchronous design borrow or are inspired by the much more developed synchronous circuits field. However, due to the difference in timing assumptions of the two types of circuits the techniques are not completely transferable. We describe why some of the solutions from synchronous circuits cannot be directly applied to asynchronous circuits.\\

%- used in the early days for space and flight, and systems that CANNOT go offline/errors like stocks and banking.
%- different techniques... explain a bit about each
%-hardware, software, hybrid, 
%-detection vs correction
%- ECC/ parity checking (needs software/ controller) (Mitra's method)
%-TMR is still super important (possibly why TMR is not viable for async)
In the early days, there were certain systems for which eliminating faults and errors were very important. These include space and aviation control systems where due to the altitude there are more radiation and thus more faults. In addition, faults that cause error is very detrimental to these systems and may cause failures leading to loss of life. Other examples of systems requiring reliable circuits include banking and stock market applications where the system must function without error for hours at a time and in case of any error may have devastating consequences, such as a bit flip of the most significant bit in a money transaction. \\

Solutions for eliminating these transient fault-induced errors include physical, hardware, software or a mix of the previous. Some commonly used techniques are transistor sizing, circuit hardening, hardware duplication, and time redundancy \cite{Mitra_08_softerrors}. Transistor sizing is a physics-based technique and does not reduce as much soft errors as the other techniques. Circuit hardening is on the physics level, where it uses intrinsic properties such as resistive or capacitive hardening, or combining the transistors differently to have less effects from transient faults. (?) Hardware duplication, is a common technique that tries to eliminate soft errors on the circuit level. This is achieved by having multiple duplicates of a circuit and often comparing the value from each duplicate to find and eliminate errors. For example, Triple Modular Redundancy (TMR) has been used since the beginning of the field and is still a good solution for eliminating single soft errors. The advantages are that it is able to eliminate nearly all soft errors but also due to the duplication, the area for each circuit and additional power increases substantially (correlates to the number of additional duplicates). Time redundancy techniques re-use the same circuitry to compute values multiple times to ensure there is no error. This technique is not completely hardware based and requires software to manage re-using the same circuity with the same inputs over time. The area of the circuit does not change but the speed of the circuit worsens substantially. \\

There is also the additional distinction between error detection versus error correction. In general, it is easier to detect an error than to correct it. For example, using hardware duplication, only 2 copies are required to detect an error but 3 copies (TMR) to correct the error. For different parts of the system there are also different strategies. For example, in memory components such as SRAM, soft error correction is usually done using error-correction codes. These codes are compared to the bits that are stored to ensure that there are no errors. They are designed to handle a predefined maximum number of errors. For example, Hamming codes where each codeword are at least a distance of 3 apart can handle one error. This method is not purely hardware based and a controller system and software are used to match the codes and detect and correct any errors. For a control circuit, there are also a variety of strategies. Again, one could make use of error correcting codes or parity bits to detect or correct errors in combinational logic \cite{McCluskey_99}. This uses a combination of hardware and software. A completely software approach is to run a program on some input, then run the program again on a transformed version of the input where the output can be mapped back to the original output \cite{Mitra_softwareerror}. If the output and the transformed output do not match then the fault is detected. Lastly, for hardware only methods, even now TMR is a commonly used method because of its robustness and ensures error correction. \\
%However these cannot be used for latches or flip flops. In these circuits, some of these techniques only detect errors but do not correct them.\\

Triple Modular Redundancy is a technique where each combinational logic unit is copied three times and then fed into a latch that performs a majority voting operation. That is, if at least 2 out of the 3 signals agree on a value, the output becomes that value. Additionally, to counter possible faults on the single output wire, the majority vote unit can also be copied three times and its output wire fed appropriately as input to the three copies of the downstream combinational logic. This would resolve both transient and permanent single faults. However, the downside is that the circuit needs to be copied three times so the area is three times larger and the power consumption as well. In addition, there must be many adaptations for this to work in asynchronous circuits as the signals can arrive in different orders. We give an example of why TMR does not work in an asynchronous circuit as it does in a synchronous circuit below. We assume that in the original circuit that each logic block has some input wires and an output wire. In the TMR scheme, each logic block is triplicated and to remove single points of failure, we also triplicate the majority voter that the three copies of the output wire feeds into. There are also three distinct copies of each input signal that correspond to outputs from the three voters. This is the basic scheme that synchronous circuits with TMR uses. Because this circuit is asynchronous, we assume that each logic block is composed of gates that has some arbitrary delay that is independent of other gates. This is true under both delay assumptions of speed independent and delay insensitive circuits. We ignore whether a fault can be in the voter for now. 
\begin{center}
\includegraphics[width=\textwidth]{tmrcounterex}
\end{center}
\captionof{figure}{TMR asynchronous circuit example with fault}
Initially all output wire values are `0's. Then since the input signals are distinct copies of each other, it is possible for one set to finish transitions (new) before the other two sets and the first logic block transitions to a `1'. Then a transient fault arrives on the output of the second logic block which transitions to a `1' as well. The majority voter would then transition to a `1'. However, the upset from transient faults usually happen within a short time period and the old inputs eventually force the output of the second logic block back to `0'. Then the majority voter would transition to a `0' so that causes a hazard on the output of the majority voter. %depends on if the logic gate was combinational logic.... which I guess we can assume it is
%NEEDS FIXING AND CHECKING
It is also possible for the `1' from the majority voter to feed into other logic units which is effectively an early transition. That early transition will then put the circuit into states that are not specified by the original state graph and thus the next states cannot be predicted.
\\

%also included detection schemes because there is so few works in the literature...
David and Ginosar (1995) prove that a specific type of asynchronous circuit called self-timed combinational logic is self-checking under stuck-at-faults \cite{self_timed}. (Stuck-at-faults are permanent faults where a wire value is stuck at 0 or 1) That is if there is a single stuck-at-fault, the circuit will either stall or give an illegal output so that faults can be detected. Each input and output is encoded by 2 bits also known as dual rail encoding. 10 represents 0, 01 represents 1, and 00 represents undefined or the system is not yet ready. 11 is an illegal output that only occurs under certain faults. Each logic unit is composed of four subunits, three of which control the inputs and outputs and includes many C-elements. The fourth one implements the combinational logic. The combinational logic requires the implementation to be such that each output bit is monotonous given the input bits. A further restriction is that the combinational logic is composed of only AND and OR gates.
\\

Verdel and Makris proposes that simply porting the duplication strategy for error detection from synchronous circuits does not work for asynchronous circuits \cite{async_dup_ced}. This is because the duplication strategy relies on both the original and duplicate to output the same value. Some modifications that they propose to make the duplication error detection strategy work is to assume additional timing requirements. That both copies should output the same value within a certain time determined by a pre-set delay element. Everytime one of the signals transitions to a new value, and the duplicates are mismatched the timer starts and if they stay mismatched when the timer ends an error is detected. They look at stuck-at-faults and premature firing faults. There are also cases when the timer detection does not work and for those cases, they propose using additional hardware to look at the states of all the wires to see if an illegal state was reached. They present an example of a control circuit for data computations. The main restrictions of this method are the additional timing assumptions they make about two copies of a circuit, and that the extra hardware components are assumed to be fault-free. Also, it requires an additional comparison unit for each pair of signals which may be costly. Only one copy of the duplicates produces output signals that interfaces with the other units so that if there is an error on that signal it is not detected until some time after the error has propagated.
\\

Almukhaizim and Sinanoglu (2007) propose a Triple Modular Redundancy (TMR) based method of solving transient errors but requires a hazard-free majority voter. Transferring the TMR method that solves faults for the synchronous circuits to asynchronous requires the change that 2/3 copies agreeing is not good enough but 3/3 is needed to propagate a signal. This is due to the asynchronous property where signals can change at any given time and thus one signal changing plus a transient error may cause the output wire to transition prematurely. For permanent errors there is an additional problem of deadlock and this configuration does not solve it completely. \\
\\
In the paper by Manohar and Peng (2005), they add in new circuitry that solves faults for a specific asynchronous adder circuit \cite{PengManohar_asyncadder}. The faults that they consider are fabrication defects, permanent faults and unrecoverable failures. New nodes and edges are added to make it k fault-tolerant. They detect when there is a deadlock at which time it reconfigures the circuit by trial and error of all the possible combinations until the circuit is no longer deadlocked. \\

Naqvi (2014) shows how to solve transient faults on muller pipelines \cite{Naqvi_mullerpipeline}. Muller pipelines make use of handshaking protocols. The assumptions they make are that only one fault occurs per handshake cycle. They used formal verification to show that it works for one pipeline component. \\
\\
Jang and Martin (2005)’s paper considers the problem of single SEUs in QDI circuits \cite{JangMartin_SEUQDI}. In the original functioning circuit, an SEU may cause a) a transition to a deadlocked state, an illegal state with no further possible transitions, b) abnormal, a transition to another state in the state graph and skip or adds certain transitions/signalling, c) tolerant where it transitions to an illegal state but the PRs are unaffected and eventually resumes to a valid state in the state graph. They solve this problem by duplicating the circuit and appending C-elements to the end of each of the original wires. In addition to duplicating the circuit they change the production rule set so that if the rule is about wire $X$ is true in the original circuit, in the new circuit the rule is $X^1$ and $X^2$ both have to be true. For example
\begin{align*}
B_u (...,x,...)\mapsto z\uparrow \\
B_d (...,\lnot x,...)\mapsto z\downarrow
\end{align*}
after duplicating the circuit and adding in C-elements becomes:
\begin{align*}
B_u^{double} (...,x_a\wedge x_b,...) & \mapsto z'_a\uparrow, z'_b\uparrow\\
B_d^{double} (...,\lnot x_a\wedge \lnot x_b,...) & \mapsto z'_a\downarrow, z'_b\downarrow\\
z'_a \wedge z'_b& \mapsto z_a\uparrow, z_b\uparrow\\
\lnot z'_a\wedge \lnot z'_b& \mapsto z_a\downarrow, z_b\downarrow
\end{align*}
The last two production rules correspond to the additional C-elements added. 
They prove that with this new circuit, single SEU errors do not cause transitions to deadlocked states, nor abnormal transitions and thus are corrected. Given the transistor-based QDI assumptions, any changes to the production rules are implemented at the transistor level and assumed to not add any delays. In addition, QDI logic is implemented in sum-of-products form and the production rules are excited only in the appropriate non-illegal states. Faults are assumed to occur on the input/output wires as well as the new wires that feed into the C-elements. Another assumption is every loop in the control circuit is at least length 3 to prevent deadlock. %not sure about this, may only be relevant for when they don't double the inputs
\\

The circuit transformation that Jang and Martin propose work for their assumptions and specific implementation of QDI circuits at the transistor level. However, it is not reasonable to assume that any PR can be implementation with transistors with no delay. Also, for a simple 2-element inverter circuit their transformation also can end up in a deadlocked state after an SEU. After the duplication and addition of C-element. 
\begin{center}
\includegraphics[width=.7\textwidth]{transfex}
\end{center}
Then if a 0->1 SEU occurs at $c_1^B$, it is possible for this circuit to deadlock.
\begin{center}
\includegraphics[width=.7\textwidth]{transfexc}
\end{center}
%fix later... it's technically that a simple duplicate and C-element scheme does not work

The approach that we use to make circuits fault-tolerant will be described in the next chaper. Some of the previous works makes additional assumptions such as the checking circuit is fault free or that only one fault can occur on an arbitrarily large transistor gate. We allow for standard two-input gates and assume that all gates in our model are subject to faults. Our fault-tolerant transformation procedure is for a class of circuits, semi-modular control circuits, which includes many circuits.

\chapter{Asynchronous Circuit Concepts \& Definitions}
\section{Assumptions:  Boolean signals}
We make the assumption that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the gates and C-elements are excited, and their outputs transition, we assume that transition is instantaneous and that after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.

>>>>

    \title{Transient Fault Tolerant Asynchronous Circuits}
    \author{Zoey Zhou}
    \dept{Electrical Engineering}
    \principaladvisor{David Dill}
    \firstreader{Subhasish Mitra}
    \secondreader{Clark Barrett}

@@

\prefacesection{Abstract}
Asynchronous digital circuits are circuits that don’t have clocks. They sometimes have advantages including speed, lower power consumption, and relative insensitivity to environmental conditions such as temperature. Transient faults, which are changes in the circuit that do not permanently damage the circuit, are a growing problem in digital circuits. Transient faults are mainly due to radiation effects such as Single Event Upsets (SEUs). These faults occur on the ground and are exacerbated in space.

In this talk, I will present an overview of asynchronous circuits and in particular speed-independent control circuits, and our model for transient faults. I will then present a fault tolerant transformation of the circuit based on duplication. I will discuss how to define fault-tolerance formally. Then I will show using logic and formal methods why the transformed circuit is fault tolerant to single transient faults.
\newpage

%% one can also have a prefacesection that is a Preface instead of
%% Acknowledgements.   The thematic purpose is the same (thanks).
%    \prefacesection{Acknowledgements}
 %       I would like to thank Professor David Dill.

%% afterpreface produces a table of contents and any other tables
%% wanted. At the end pagenumbering changes from roman to arabic and
%% is restarted
\tablespagefalse
    \afterpreface

@@

%However, because the signals can arrive at the input at any time, the design is more complicated to ensure that the entire circuit functions correctly.  From a high level view the circuits are designed so that the signals progress irrespective of the delays.  This property is called Delay Insensitivity or Speed Independence depending on the underlying delay assumptions such as the maximum length of delays and on which components.  Even if this is designed correctly, when implemented there can still be hazards.  A hazard is when a certain sequence of delays of the circuit cause unwanted signal transitions to occur.  This may then propogate to other parts of the circuits and cause unwanted behavior.  When trying to eliminate hazards, the underlying delay assumptions are important as well as the implementation details such as gates and transistors. Given the same design under one implementation the circuit may have hazards while under another implemention the circuit will be hazard-free.  Some of the different delay assumptions that could be made are that the delays only occur at gates, or delays only occur on the wires or the delays occur on both gates and wires \cite{myers_book_2004}. Furthermore the bounds on the delay can vary as well. The most stringent case is if we assume that the delay can be arbitrarily long and takes on a value anywhere in the interval $(0,\infty)$ \cite{myers_book_2004}.  Circuits built under the unbounded gates and wires delay model are called Delay Insensitive circuits.  We will not use this delay model because the class of Delay Insensitive circuits is very restricted.  In particular one can only use buffers, inverters, and Muller C-elements to build Delay Insensitive circuits \cite{Martin_1990_DI} \cite{Martin1986_DI}.\\ %2nd reference not necessary

The asynchronous delay model that we consider is the unbounded gate delay model.  Circuits that can function correctly under arbitrary gate delays are called Speed Independent circuits. The unbounded gate delay model assumes that delays only appear on the gates and the delays can be arbitrarily long in the interval $(0,\infty)$.  The wire delays are negligible.  In addition we assume all forks, which is splitting a signal into multiple branches, do not have any time delays between each of its branches.  These assumptions are reasonable because the delays on the wires only change the operation of a circuit if the wire has forks.  (Otherwise you can lump the wire delay with the previous gate)  When a wire has forks, it is usually true in implementation that the delay between different branches of the fork is less than the delay for gates so that they arrive at the next gates at the same time from the perspective of the relevant gates.  This is also termed isochronic forks.  We focus on the unbounded gate delay model because the delay being unbounded is a pretty harsh requirement and would allow more circuits to work in a real life setting. \\ %how if we assume both gates and wires have delay cannot create many SI circuits, and also Muller is one of the broader classes that we can use

@@

The standard C-implementation is a gate-level implementation.  Each wire from the specification is implemented as a C-element with non-looping cascades of gates as the C-element's two inputs.  That is, each of the two inputs to the C-element is itself an output of a cascade of gates.  The C-element operates like a typical C-element where if the two inputs are `1', then the output is `1'; and when the two inputs are `0', the output is `0'; and when the two inputs have different values, the output value is held.  Additionally, Myers shows a method where the gate logic for the inputs to the C-element can always be re-implemented with two-input gates \cite{myers_book_2004}.  Since the C-element itself is also a two-input circuit element, then the entire circuit is composed of two-input circuit elements.  In such an implementation, each gate or C-element is assumed to have its own delay that is arbitrary in length.  Thus, this gate-level implementation allows for more generic design.
%explain sop?
\begin{comment}
\subsection{Quasi Delay Insensitive Circuits}
%QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits. Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays. In particular each branch of a fork will have different delays and can be treated as separate wires. One result about Delay Insensitive circuits is that it is a very restricted class. Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed. Note that in the case where all forks are isochronic, then this becomes the same as speed independent circuits. \\
The gates of a QDI circuit is formed from Production Rules (PRs). For the single output case, a gate takes a set of variables as inputs and outputs a variable. Again, a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$. Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards. Each guard is a boolean condition formed by some combination of the input variables. When a guard is evaluated to be $True$ then the transition is activated. When it is $False$ then the corresponding transition does not happen. Note that similar to the generalized C-elements, both guards cannot be active at the same time. In addition, it is assumed that each guard is in a sum of products of the input variables. The PRs of a circuit are evaluated all at the same time. At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$. This is due to the assumption of the arbitrary delay of a gate. Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allows any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

Another important property of QDI circuits is \textit{stability}. A guard is stable if it evaluates to False after the transition of the output variable is complete. This eliminates any hazards that may arise when a guard is True and before the output has fully transitioned, the guard then evaluates to False and the transition is terminated. This may cause a blip in the voltage, an incomplete transition, where the circuit may interpret it as either no transition or two consecutive transitions. One additional assumption is that the PRs cannot be \textit{self-invalidating}. That is, when a transition on an output occurs, it cannot immediately falsify the guard. With this assumption, it can be shown that the output of a gate cannot also be its own input variable. There are no self-loops. 
\end{comment}
\section{Fault Tolerant solutions}

%soft error protection <- error correcting codes for SRAM memory 
There has been significant previous work on fault-tolerant synchronous circuits.  Various methods exist that allow a general circuit to be fault-tolerant to permanent and transient faults.  However, results on asynchronous circuits are more limited.  Usually, the method is fault-tolerant to only one of permanent or transient faults, and applies to a limited category of asynchronous circuits.  Many fault-tolerant techniques in asynchronous circuits borrow from the developed synchronous circuits field.  However, due to the differences of the two types of circuits, these techniques need to be modified to work in asynchronous circuits.  The rest of this section gives an overview of some common fault-tolerance techniques for synchronous circuits, then there is a summary of the current techniques for asynchronous circuits. \\

%- used in the early days for space and flight, and systems that CANNOT go offline/errors like stocks and banking.
%- different techniques... explain a bit about each
%-hardware, software, hybrid, 
%-detection vs correction
%- ECC/ parity checking (needs software/ controller) (Mitra's method)
%-TMR is still super important (possibly why TMR is not viable for async)

%In the early days, there were certain systems for which eliminating faults and errors were very important. These include space and aviation control systems where due to the altitude there are more radiation and thus more faults. In addition, faults that cause error is very detrimental to these systems and may cause failures leading to loss of life. Other examples of systems requiring reliable circuits include banking and stock market applications where the system must function without error for hours at a time and in case of any error may have devastating consequences, such as a bit flip of the most significant bit in a money transaction. \\

In synchronous circuits, solutions for eliminating transient faults include physical, hardware, software or a mix of the previous.  Some commonly used techniques are transistor sizing, circuit hardening, hardware duplication, and time redundancy \cite{Mitra_08_softerrors}.  Transistor sizing is a physics-based technique and does not reduce transient faults as much as the other techniques. Circuit hardening is also on the physics level, it uses intrinsic properties such as resistive or capacitive hardening, or placing the transistors differently to have less effects from transient faults.  Redundancy techniques work for both transient and permanent faults.  
The first type of redundancy is hardware duplication, which eliminates faults on the circuit level. This is achieved by duplicating a circuit multiple times and using a majority voter to eliminate the faults.  For example, Triple Modular Redundancy (TMR) eliminates single faults.  However, the disadvantage is that the power and area of the circuit increases substantially.  Another type of redundancy is time redundancy.  This technique re-use the same circuitry to compute values multiple times and compare the results. This technique requires software to computing with the same circuity multiple times. In this case, the area of the circuit does not change but the speed of the circuit worsens substantially. \\

There is also the distinction between error detection versus error correction. In general, it is easier to detect an error than to correct it. For example, using hardware duplication, only 2 duplicates are required to detect an error but 3 duplicates (TMR) are required to correct the error.  There are also different strategies for different types of circuits. For example, in memory components such as SRAM, transient fault induced error is usually corrected using error-correction codes.  Error-correcting codes can also be used for a control circuit to detect or correct errors in combinational logic \cite{McCluskey_99}.  This uses a combination of hardware and software.  Another approach is software-based and inputs the original and a transformed version of the inputs to the circuit.  The transformed version is designed such that if the circuit is fault-free, the outputs map back to the original outputs \cite{Mitra_softwareerror}.  Finally, for hardware only approaches, TMR is a commonly used method because of its robustness and fault correction. \\
%However these cannot be used for latches or flip flops. In these circuits, some of these techniques only detect errors but do not correct them.\\

Triple Modular Redundancy is a technique where each combinational logic unit is copied three times and then fed into a latch that performs a majority voting operation. If at least 2 out of the 3 signals agree on a value, the output becomes that value.  Additionally, to counter possible faults on the output wire, the majority voting unit can also be copied three times and each output wire fed as input to one of the three copies of the downstream combinational logic. TMR eliminates both transient and permanent single faults.  However, the downside is that the circuit needs to be copied three times, so the circuit area and power consumption is three times as much.  Although this technique is powerful in synchronous circuits, it does not work in asynchronous circuits as is.  We give an example of why TMR does not work in an asynchronous circuit below. We assume that in the original circuit, each logic block has some input wires and an output wire. In the TMR scheme, each logic block and majority voter is triplicated. There are also three distinct copies of each input signal that correspond to outputs from the three voters. Because this circuit is asynchronous, we assume that each logic block is composed of gates that has some arbitrary delay that is independent of other gates.   
\begin{center}
\includegraphics[width=\textwidth]{tmrcounterex}
\end{center}
\captionof{figure}{TMR asynchronous circuit example with fault}
Initially, all output wire values are `0's. Then, since the input signals are distinct copies of each other, it is possible for one set to finish transitions (new) before the other two sets and the first logic block transitions to a `1'. Then a transient fault arrives on the output of the second logic block which transitions to a `1' as well. The majority voter would then transition to a `1'. However, the old inputs can force the output of the second logic block back to `0'. Then the majority voter would transition to a `0' which causes a hazard on the output of the majority voter. %depends on if the logic gate was combinational logic.... which I guess we can assume it is
%NEEDS FIXING AND CHECKING
%It is also possible for the `1' from the majority voter to feed into other logic units which is effectively an early transition. That early transition will then put the circuit into states that are not specified by the original state graph and thus the next states cannot be predicted.
Almukhaizim and Sinanoglu (2007) solves this problem by requiring a hazard-free majority voter.  Instead of a majority voter, this design requires that all three signals match to propagate the signal.  However, with this new modification, TMR deadlocks when there are permanent faults.  Thus, this design is only for transient faults in asynchronous circuits \\

%also included detection schemes because there is so few works in the literature...
There are also permanent fault solutions in asynchronous circuits.  
David and Ginosar (1995) show that faults can always be detected when there is a permanent stuck-at fault.  This is restricted to self-timed combinational logic composed of only AND and OR gates.  
Manohar and Peng (2005) designed fault-tolerant asynchronous adders that can handle up to $k$ permanent faults \cite{PengManohar_asyncadder}.  This is achieved by adding new circuitry and restarting different parts of the circuit when it deadlocks due to the faults.  Verdel and Makris proposed a duplication method to detect permanent faults \cite{async_dup_ced}.  They assumed that both duplicates output a new value within a certain amount of time of each other.  For each pair of signals, they use a timer to detect when that timing is violated, and the circuit is in deadlock.  However, this timer method does not catch all faults and they propose to add circuitry to detect when additional illegal states are reached.    
\\

Naqvi (2014) gives a fault-tolerant solution to transient faults on muller pipelines \cite{Naqvi_mullerpipeline}.  This is a duplication and design based technique that allows for one fault per handshake cycle.  Formal verification is used to show that one pipeline component can be made fault-tolerant. \\

Jang and Martin (2005)’s work is the closest existing work to ours. Their design makes Quasi-Delay-Insensitive (QDI) circuits fault-tolerant to single SEUs \cite{JangMartin_SEUQDI}. QDI uses a similar delay assumption as Speed-Independent circuits and can also model control circuits. However, QDI circuits are synthesized differently from the methods outlined for Speed-Independent circuits. Instead, the high-level specifications are computed into two Production Rules (PRs) for each gate. The PRs are like the set and reset logic for generalized C-elements. For each gate, there is a PR for when the output transitions to `1', and another PR for when the output transitions to `0'. For example, if a gate takes a set of wires as inputs and outputs a wire $x$, and a transition from $x=0$ to $x=1$ is denoted by $x\uparrow$, while a transition from $x=1$ to $x=0$ is denoted by $x\downarrow$, then the two PRs associated with the gate is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}

The Boolean expressions $B_u$ and $B_d$ are called guards, and take wires as inputs.  Also like the generalized C-element, each gate is implemented at the transistor level and both guards evaluate at the same time.\\

With this construction in mind, we look to their design.  Their solution is duplication based and a layer of C-elements is added to the output of each of the original wires. In addition to duplicating the circuit, the PRs are modified so that if the PR requires wire $X$ to be true in the original circuit, then in the fault-tolerant circuit, the new PR requires both $X^1$ and $X^2$ to be true. For example:
\begin{align*}
B_u (...,x,...)\mapsto z\uparrow \\
B_d (...,\lnot x,...)\mapsto z\downarrow
\end{align*}
after duplicating the circuit and adding in C-elements becomes:
\begin{align*}
B_u^{double} (...,x_a\wedge x_b,...) & \mapsto z'_a\uparrow, z'_b\uparrow\\
B_d^{double} (...,\lnot x_a\wedge \lnot x_b,...) & \mapsto z'_a\downarrow, z'_b\downarrow\\
z'_a \wedge z'_b& \mapsto z_a\uparrow, z_b\uparrow\\
\lnot z'_a\wedge \lnot z'_b& \mapsto z_a\downarrow, z_b\downarrow
\end{align*}
The last two production rules correspond to the additional C-elements. 
With this circuit transformation, single SEU errors do not cause transitions to deadlocked states, nor abnormal transitions and thus are corrected. Given the transistor-based QDI assumptions, any changes to the production rules are implemented at the transistor level and assumed to not add any delays.  Faults are assumed to occur on the input/output wires.  It is also assumed that every loop in the control circuit is at least length 3 to prevent deadlock. 
\\

Jang and Martin’s circuit transformation works under certain assumptions. However, the solution is not suitable for cases where those assumptions are not satisfied, such as if it is impossible to implement the gate at the transistor level with no delay, or if there are loops of length 2 in the circuit.  For example, a simple two element inverter circuit can end up in a deadlocked state after an SEU with this transformation. Then if a 0->1 SEU occurs at $c_1^B$, it is possible for this circuit to deadlock.
\begin{center}
\includegraphics[width=.7\textwidth]{transfex}
\end{center}

\begin{center}
\includegraphics[width=.7\textwidth]{transfexc}
\end{center}
%fix later... it's technically that a simple duplicate and C-element scheme does not work

We learn from previous work that it is difficult to design fault-tolerant asynchronous circuits.  Many of the previous works make additional assumptions about where the fault can occur or the timing of delays or restrict the type of circuits.  We add to this body of work with a new transient fault-tolerant design for a general class of asynchronous circuits, semi-modular control circuits.  We assume the circuit is composed of two-input gates and that each gate has variable delay and is also subject to faults.

\chapter{Asynchronous Circuit Concepts and Definitions}
\section{Assumptions:  Boolean signals}
This work assumes that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the gates and C-elements are excited and their outputs transition, we assume that the transition is instantaneous and that, after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.
